---
title: "Linear Regression"
author: "My An Huynh"
date: "2024-11-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Linear models

y = beta0 + beta1x1 + ... + betapxip + ei,ei ~ (0,sigma^2)

Preictors:
Outcome is continuous; predictors can be anything 
Continuous predictors are added directly
Categorical predictors require dummy indicator variables 
- For each non-reference group, a binary(0/1) variable indicating group membership for each subject is created and used in the model. 

Testing
- For a single regression coefficient, you can construct a test statistic using t = (beta_hat - beta) / SE(beta) where beta is the null and beta_hat is the estimate. For large samples, this has a standard normal distribution. To test multiple coefficients (i.e those arising from the inclusion of a categorical variable with several predictors) but can use an F-test / ANOVA. 

Diagnostics
- Many model assumptions (constant variance, model specification) can be examined using residuals
  - Look at overall distribution (centered at 0? skewed?)
  - Look at residuals vs predictors (any non-linearity? trends? non-constant residual variance?)
  - Are there structures in residuals? Have you fit the right trendline? Is there low variance vs high variance on different sides of the fit? 
  - Departures from these assumptions means that the statistical tests you have done then downstream things/results are also not right. 
  
Linear models in R 
- lm for linear models
- glm for generalized linear models 

Arguments include
- formula: y (output) ~ x1 + x2 (list of predictors)
- data

Output is complex --> Use broom 


```{r}
library(tidyverse)
library(p8105.datasets)
```

```{r}
data("nyc_airbnb")

nyc_airbnb  = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood
  ) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type)
```


# Fit a model 

```{r}
nyc_airbnb |> 
  ggplot(aes(x = stars, y = price, color = borough)) +
  geom_point()
```

Let's fit a model we care about. 
Fit price against stars and borough 

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

Let's look at the result 
```{r eval = FALSE}
fit
summary(fit)
summary(fit)$coef
coef(fit)

```

Let's look at the result better 
broom::glance() gives you some of the more useful stats in a tibble
broom::tidy() gives a whole df. 
stars is the reference group. For every 1 unit increase in stars (implicit neighborhood = the bronx), we expect the price of the room to go up 32 dollars keep the borough fixed.
```{r}
broom::glance(fit)

broom::tidy(fit) |> 
  select(-std.error, -statistic) |> 
  mutate(
    term = str_replace(term, "borough", "Borough: ")
  ) |> 
  knitr::kable(digits = 3)
```

## Be in control of factors
Factor borough and room type based on frequency
```{r}
nyc_airbnb = 
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type)
  )
```

Make a plot. Now the x-axis is ordered by the frequency 
```{r}
nyc_airbnb |> 
  ggplot(aes(x = stars, y = price, color = borough)) +
  geom_point()
```

Refit model. Now the reference group is Manhattan. It hasn't substantially changed what is happening. 

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)

broom::tidy(fit)
```

